---
title: "Analysis of the global temperature development at country level"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

For this analysis we are using the Climate Change: Earth Surface Temperature Data from kaggle.
This data set can be found here: 
https://www.kaggle.com/berkeleyearth/climate-change-earth-surface-temperature-data#GlobalLandTemperaturesByCountry.csv

The goal is to show how the fxtract package can support climate and economical analysis.

### Load required packages
```{r, message=FALSE}
library(fxtract)
library(lubridate)
library(tidyverse)
library(stringr)
```

### Read in data
```{r}
df_glob_temp = read.csv("GlobalLandTemperaturesByCountry.csv")
str(df_glob_temp)
```

### Data preprocessing
For our analysis we consider only years since 1900. 
The main reason for this is that in recent years we have had many missing values for some countries.
```{r}
df_glob_temp$dt = as.Date(df_glob_temp$dt)
df1900_raw = df_glob_temp %>% filter(dt >= "1900-01-01")
```

Removal of duplicates and some smaller regions.
```{r}
df1900 = df1900_raw %>% filter(!Country %in% c('Denmark', 'Antarctica', 'France', 'Europe', 'Netherlands',
  'United Kingdom', 'South America', 'Ã…land', 'Africa',  'Asia', 'Baker Island', 'Curaçao', 'Kingman Reef', 'North America', 'Oceania', 
  'Palmyra Atoll', 'Saint Barthélemy', 'Saint Martin', 'Virgin Islands'))
df1900$Country = gsub("\\(Europe)", "", df1900$Country)
```

For an annual view we would like to have the year in a separate column.
```{r}
df1900$year = year(df1900$dt)
```


## fxtract

### Create our user defined-functions
```{r, message=FALSE}
temp_stats = function(data) {
  allYears = unique(data$year)
  data = data %>% group_by(year) %>%
    summarise(
      mean = mean(AverageTemperature, na.rm = T),
      sd = sd(AverageTemperature, na.rm = T),
      min = min(AverageTemperature, na.rm = T),
      max = max(AverageTemperature, na.rm = T)
    )
  data = data %>% as.data.frame()
  res = c(mean = data$mean,
     sd = data$sd,
     min = data$min,
     max = data$max)
  allYears = unique(df1900$year)
  newnames = c(paste0("mean_", allYears), 
    paste0("sd_", allYears), 
    paste0("min_", allYears), 
    paste0("max_", allYears)) 
  names(res) = newnames
  res
}
```

### Setup and calculation
```{r, message=FALSE, results='hide', warning=FALSE}
xtractor = Xtractor$new("xtractor")
xtractor$n_cores = 2
xtractor$add_data(df1900, group_by = "Country")
xtractor$add_feature(temp_stats)
xtractor$calc_features()
```

### Results
```{r, results='asis', message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
res = xtractor$results %>% gather(key = "key", value = "value", -Country) %>% 
  separate(key, c("key", "year")) %>% select(Country, year, key, value) %>% 
  as.data.frame()
res$value[is.infinite(res$value)] <- NA 
res %>% slice(1:20) %>% kable(col.names = c("Country", "Year", "Key", "Value")) %>% 
    kable_styling() %>%
    scroll_box(width = "100%", height = "400px")

```

### Visualization
```{r, echo=FALSE}
df_sub = res %>% filter(Country %in% c("Argentina", "Congo", "Germany", "United States", "Australia"),
  key == "mean")
df_sub$year = as.numeric(df_sub$year)
ggplot(df_sub, aes(x=year, y = value, col = Country)) + geom_point() +
  geom_smooth(method = "loess") +
  scale_x_continuous(breaks = seq(1900, 2015, by = 20)) +
  ggtitle("Development of average global temperature") + 
  labs(x = "Year", y = "Average temperature [in °C]") +
  theme_bw()
```


### Linear Regression with fxtract

#### User defined functions

```{r}
limo_slope = function(data){
  lin_model = lm(temp ~ year, data = data)
  slope = lin_model$coefficients[[2]]
  c("slope" = slope)
}
```


#### Setup and calculation

```{r, message=FALSE, results='hide', warning=FALSE}

df = res %>% filter(key == "mean") %>% select(temp = value, everything())
df$year = as.numeric(df$year)
xtractor2 = Xtractor$new("xtractor2")
xtractor2$n_cores = 2
xtractor2$add_data(df, group_by = "Country")
xtractor2$add_feature(limo_slope)
xtractor2$calc_features()
res2 = xtractor2$results

```

#### Results
```{r, echo=FALSE}
res2 %>% slice(1:20) %>% kable(col.names = c("Country", "Slope")) %>% 
    kable_styling()  %>%
    scroll_box(width = "100%", height = "400px")
```

<br/><br/>

#### Visualization

For a better overview, we can then plot our results in a map.
```{r, message=FALSE,warning=FALSE, echo=F}
library(plotly)
library(countrycode)

df2 = res2
df2$code = countrycode(df2$Country, 'country.name', 'iso3c')

l <- list(color = toRGB("grey"), width = 0.5)
# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

p <- plot_geo(df2) %>%
  add_trace(
    z = ~slope, color = ~slope, colors = 'Reds',
    text = ~Country, locations = ~code, marker = list(line = l)
  ) %>%
  colorbar(title = 'Temperature change', tickprefix = '°C ') %>%
  layout(
    title = 'Average temperature change per year since 1900',
    geo = g
  )
p
```




